---
title: Will students who have high GPA in high school get high GPA in the first year
  college?
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Interpretation and Implementation

```{r}
#primary objective
#set up
#install.packages("ggplot2")
library(ggplot2)

#load the data
student_info<-read.csv("FirstYearGPA.csv",header=T)

#set up variables
#set of high school GPA of students
x<-student_info$HSGPA
#set of first year college GPA of students
y<-student_info$GPA

#student whose GPA scores above or equal 3.7 in high school
HSGPA_greater_than_critical_value<-student_info[student_info$HSGPA>=3.7,]


#student whose GPA scores below 3.7 in high school
HSGPA_less_than_critical_value<-student_info[student_info$HSGPA<3.7,]


#college GPA>=3.7 and high school GPA >=3.7
group_A<-student_info[(student_info$HSGPA>=3.7&student_info$GPA>=3.7),]

#college GPA>3.7 and high school GPA <3.7
group_B<-student_info[student_info$HSGPA<3.7&student_info$GPA>=3.7,]

#college GPA<3.7 and high school GPA >=3.7
group_C<-student_info[student_info$HSGPA>=3.7&student_info$GPA<3.7,]

#college GPA<3.7 and high school GPA <3.7
group_D<-student_info[student_info$HSGPA<3.7&student_info$GPA<3.7,]


#Graph plotting
#scatter plot
graph_GPA<-ggplot(student_info,aes(x=HSGPA,y=GPA))+geom_point(size=0.35,color="forestgreen")+geom_smooth(size=0.4,method="lm",color="red",se=F)+labs(x="High school GPA",y="First year college GPA",title="First year college GPA versus High school GPA")+theme(axis.text=element_text(size=11),plot.title=element_text(size=11,hjust = 0.5))

graph_GPA

#barchart 
#set up
group_name<-c("GPA>=3.7&HSGPA>=3.7","GPA>=3.7&HSGPA<3.7","GPA<3.3&HSGPA>=3.7","GPA<3.3&HSGPA <3.7")
group_size<-c(length(group_A$GPA),length(group_B$GPA),length(group_C$GPA),length(group_D$GPA))
group_data<-data.frame(group_name,group_size)

  

ggplot(data=group_data,aes(x=group_name,y=group_size,fill=group_name))+geom_bar(stat="identity",width=0.5,color="black",size=1)+theme_minimal()+labs(x="Group of students",y="Number of student",title=" GPA comparsion between high school and college")+theme(axis.text=element_text(size=4),plot.title=element_text(size=9,hjust = 0.5))

#Pearson’s Correlation Coefficient

size_of_sample<-length(x)

#mean of sample X
mean_x<-mean(x)

#mean of sample Y
mean_y<-mean(y)

#calculation for Pearson’s Correlation Coefficient according to the formula shown in the appendix of the report
numerator<-sum((x-mean_x)*(y-mean_y))
denominator_1<-(sqrt(sum(((x-mean_x)^2))))
denominator_2<-(sqrt(sum(((y-mean_y)^2))))

 
r<-numerator/(denominator_1*denominator_2)

r

#secondary objective
#set up
#load the data
data<- read.csv("FirstYearGPA.csv",header=T)
#set up variables
satvnc_1<- data$SATV
satvnc_2<- data$SATM
#sumsat=sum of SAT score
sumsat<- satvnc_1 + satvnc_2
#highgpa=high school GPA
highgpa<- data$HSGPA
#unigpa=first year college GPA
unigpa<- data$GPA
#familyedu=family education level
familyedu<- data$FirstGen

#boxplot:family education level and sumsat
boxplot(sumsat~familyedu, xlab="1= student is the first in her or his family to attend college,
0=otherwise")

#boxplot:family education level and high school GPA
boxplot(highgpa~familyedu,xlab="1= student is the first in her or his family to attend college,
0=otherwise")

#boxplot:family education level and first year college GPA
boxplot(unigpa~familyedu,xlab="1= student is the first in her or his family to attend college,
0=otherwise")

#correlation coefficient:family education level and sumsat
cor(familyedu,sumsat)

#correlation coefficient:family education level and high school GPA
cor(familyedu,highgpa)

#correlation coefficient:family education level and first year college GPA
cor(familyedu,unigpa)

#Tertiary objective
#read data set
data <- read.csv("FirstYearGPA.csv")

#extract the Male, HU, SS from the data set
d <- data.frame(data$Male, data$HU, data$SS)

#seperate set
m <- d$data.Male == 1
male <- d[m,]
female <- d[!m,]

#add gender
nrow(male)
nrow(female)
gender <- c(rep("Male",102), rep("Female",117))
HU <- c(male$data.HU, female$data.HU)
SS <- c(male$data.SS, female$data.SS)
df <- data.frame(gender, HU, SS)
genderm <- c(rep("Male",102))
HUm <- c(male$data.HU)
SSm <- c(male$data.SS)
dfm <- data.frame(genderm, HUm, SSm)
genderf <- c(rep("Female",117))
HUf <- c(female$data.HU)
SSf <- c(female$data.SS)
dff <- data.frame(genderf, HUf, SSf)

#graph of gender and cradit on humanity course
p1 <- hist(male$data.HU, breaks = 40)
p2 <- hist(female$data.HU, breaks = 40)
plot(p2, col = rgb(0, 0, 1, 1/2), main="Credit Hours on Humanity of Males and Females", xlab="credit hours", ylim = c(0,10))
plot(p1, col = rgb(1, 0, 0, 1/2), add = T)
abline(v = mean(male$data.HU), col="blue")
abline(v = mean(female$data.HU), col="red")
legend("topright", c("Male", "female"), fill = c(rgb(0, 0, 1, 1/2), rgb(1, 0, 0, 1/2)))

#graph of Credit Hours on Social Science course
p1 <- hist(male$data.SS, breaks = 25)
p2 <- hist(female$data.SS, breaks = 25)
plot(p2, col = rgb(0, 0, 1, 1/2), main="Credit Hours on Social Science of Males and Females", xlab="credit hours", ylim = c(0,18))
plot(p1, col = rgb(1, 0, 0, 1/2), add = T)
abline(v = mean(male$data.SS), col="blue")
abline(v = mean(female$data.SS), col="red")
legend("topright", c("Male", "female"), fill = c(rgb(0, 0, 1, 1/2), rgb(1, 0, 0, 1/2)))

#graph of Credit Hours on Social Science and Humanity of males
p1 <- hist(male$data.HU, breaks = 40)
p2 <- hist(male$data.SS, breaks = 20)
plot(p2, col = rgb(0, 0, 1, 1/2), main="Credit Hours on Social Science and Humanity of Males", xlab="credit hours", xlim = c(0,40), ylim = c(0,17))
plot(p1, col = rgb(1, 0, 0, 1/2), add = T)
abline(v = mean(male$data.SS), col="blue")
abline(v = mean(male$data.HU), col="red")
legend("topright", c("Social Science", "Humanity"), fill = c(rgb(0, 0, 1, 1/2), rgb(1, 0, 0, 1/2)))


#graph of Credit Hours on Social Science and Humanity of males
p1 <- hist(female$data.HU, breaks = 40)
p2 <- hist(female$data.SS, breaks = 25)
plot(p2, col = rgb(0, 0, 1, 1/2), main="Credit Hours on Social Science and Humanity of Females", xlab="credit hours", xlim = c(0,40), ylim = c(0,18))
plot(p1, col = rgb(1, 0, 0, 1/2), add = T)
abline(v = mean(female$data.SS), col="blue")
abline(v = mean(female$data.HU), col="red")
legend("topright", c("Social Science", "Humanity"), fill = c(rgb(0, 0, 1, 1/2), rgb(1, 0, 0, 1/2)))

#mean
mean(male$data.SS)
mean(male$data.HU)
mean(female$data.SS)
mean(female$data.HU)

#normality test
shapiro.test(male$data.SS)
shapiro.test(male$data.HU)
shapiro.test(female$data.SS)
shapiro.test(female$data.HU)

#variance test
bartlett.test(HU ~ gender, df)
bartlett.test(SS ~ gender, df)

```
document:
male: subset that contains the sample data of male students.
female: subset that contains the sample data of female students.
male$data.HU: Credit hours on humanity of male subset
female$data.HU: Credit hours on humanity of female subset
male$data.SS: Credit hours on social science of male subset
female$data.SS: Credit hours on social science of female subset

